---
title: "Code"
output: 
  html_document:
    keep_md: true  # Optionally keep intermediate markdown
    toc: true  # Table of contents, optional
    toc_float: true  # Float the table of contents for easier navigation, optional
    number_sections: true  # Number sections, optional
date: "2024-09-06"
---

```{r setup, include=FALSE}

rm(list = ls())


library(ggplot2)
library(dplyr)
library(fs)
library(readxl)
library(scales)
library(showtext)
library(lubridate)
library(zoo)
library(gridExtra)
library(tinytex)
library(tidyverse)
library(rio)
library(ggridges)
library(ggpattern)
library(ggdist)
library(ggbeeswarm)
library(ggstream)
library(ggpubr)
library(scales)
library(knitr)
library(reactable)
library(htmltools)
library(zoo)
library(quantmod)
library(PerformanceAnalytics)
library(quadprog)
library(janitor)
library(stringi)
library(car)
library(ggplot2)
library(dplyr)
library(ggdist)

knitr::opts_chunk$set(echo = TRUE)

```

```{r}
font_add_google("Open Sans")
showtext_auto()

# Define theme_reason based on the uploaded file
theme_reason <- function(base_line_size = 0.5,
                         base_family = "Open Sans",
                         base_size = 12) {
  theme_minimal(base_family = base_family,
                base_size = base_size,
                base_line_size = base_line_size) %+replace%
    theme(
      axis.ticks = element_line(),
      panel.grid = element_blank(),
      panel.grid.major.y = element_line(color = "grey92"),
      legend.position = "top"
    )
}

C1 <- "#CC0100"  # Light orange 
C2 <- "#66BB8B"  # Mid-tone orange
C3 <- "#FF6733"  # Warm orange
C4 <- "#2F456B"  # Lighter orange
C5 <- "#A69FA1"  # Pale orange
C6 <- "#F6B941"  # Yellow-orange
C7 <- "#2779CB"  # Bright orange

reason_colors <- c(C1, C2, C3, C4, C5, C6, C7)

```

```{r}
combined_data_clean <- read.csv("combined_data_clean.csv")

```


```{r}

```

# 1 The Disparity in Fiscal Years 

# 1.1 Figure 1: Number of Plans filing in each fiscal year end month

```{r}
data_2022 <- combined_data_clean %>%
  filter(fye >= as.Date("2022-01-01") & fye <= as.Date("2022-12-31"))

# Calculate the number of plans that file in each fiscal year-end month for 2022
plan_counts_2022 <- data_2022 %>%
  group_by(month) %>%
  summarise(plan_count = n_distinct(plan))  # Count distinct pension plans per month

# Create a bar chart displaying the number of plans per fiscal year-end month
Bars <- ggplot(plan_counts_2022, aes(x = as.factor(month), y = plan_count, fill = as.factor(month))) +
  geom_bar(stat = "identity", show.legend = FALSE) +  # Bar chart
  scale_fill_manual(values = reason_colors) +  # Assign the colors to each month
  labs(x = "Fiscal Year-End Month", y = "Number of Pension Plans", 
       title = "Number of Pension Plans Filing by Fiscal Year-End Month (2022)") +
  theme_reason()  # Clean theme for the chart

Bars

ggsave("outputs/Bars.png", plot = Bars, width = 16, height = 7)

```





```{r}

plan_counts_2022 <- data_2022 %>%
  group_by(month) %>%
  summarise(plan_count = n_distinct(plan))  # Count distinct pension plans per month

# Calculate total number of plans in 2022
total_plans_2022 <- sum(plan_counts_2022$plan_count)

# Calculate the percentage of plans filing in each month
plan_counts_2022 <- plan_counts_2022 %>%
  mutate(percentage = (plan_count / total_plans_2022) * 100)

# Display the table with fiscal year-end month and percentage of total plans
plan_counts_2022 %>% 
  arrange(month) %>% 
  knitr::kable(col.names = c("Fiscal Year-End Month", "Number of Plans", "Percentage of Total Plans"))


```


# 2. 2022 Returns by Fiscal Year End Month Cohort 

# 2.2. 2022: Average Returns by Month End
```{r}
data_2022_2 <- combined_data_clean %>%
  filter(fye >= as.Date("2022-01-01") & fye <= as.Date("2022-12-31"))

# Calculate the mean return for each fiscal year-end month
mean_returns_by_month <- data_2022_2 %>%
  group_by(month) %>%
  summarise(mean_return = mean(return, na.rm = TRUE)) %>%  # Calculate the mean return for each month
  mutate(mean_return = round(mean_return * 100, 2))  # Convert to percentage and round to 2 decimal places

# Display the average returns table for fiscal year 2022
mean_returns_by_month %>%
  knitr::kable(col.names = c("Fiscal Year-End Month", "Mean Return (%)"),
               caption = "Mean Returns by Fiscal Year-End Month for 2022",
               format = "latex")  # or omit `format` to be compatible with both

# Print the mean returns by month
print(mean_returns_by_month)
```

# 2.3. Average 2022 Return For all Plans

```{r}
overall_avg_return_2022 <- data_2022 %>%
  summarise(arithmetic_avg_return = mean(return, na.rm = TRUE)) %>%
  mutate(arithmetic_avg_return = round(arithmetic_avg_return * 100, 2))

# Print the overall average return for the fiscal year 2022
print(paste("Overall Average Return for FY 2022:", overall_avg_return_2022$arithmetic_avg_return, "%"))

```


# 2.4. Figure 2: Distribution of 2022 Returns by Fiscal Year End Month Cohort

```{r}
data_2022 <- combined_data_clean %>%
  filter(fye >= as.Date("2022-01-01") & fye <= as.Date("2022-12-31"))

return_plan_22 <- data_2022 %>%
  group_by(plan) %>%
  summarise(avg_return = mean(return, na.rm = TRUE),  # Calculate arithmetic mean for each plan
            month = mean(month, na.rm = TRUE))  # Get the mean fiscal year-end month for each plan

D22 <- ggplot(return_plan_22, aes(x = as.factor(round(month)), y = avg_return, fill = as.factor(round(month)))) +
  stat_dotsinterval(slab_linewidth = 0, 
                    side = "both", 
                    binwidth = NA,  # Automatically determine binwidth
                    overflow = "compress",  # Compress spacing to fit within bounds
                    .width = c(0.50),
                    show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = reason_colors) +
  labs(x = "Fiscal Year-End Months", y = "2022 Average Returns",
       caption = "Public Pension 2022 Return Distribution By FY End Months") +
  theme_reason()

D22

ggsave("outputs/D22.png", plot = D22, width = 16, height = 7)

```
# 3. Long Run Returns by Fiscal Year End Monrh Cohort 

# 3.1 Average 10 and 20 year Returns For All plans  

```{r}
return_national_fye <- combined_data_clean %>%
  filter(!is.na(mva)) %>%
  group_by(month, fye) %>%
  summarise(
    national_return = mean(return, na.rm = TRUE),  # Changed to simple mean for unweighted average
    national_ARR = mean(arr, na.rm = TRUE),  # Changed to simple mean for unweighted average
    .groups = "drop"  # Override the default grouping behavior
  ) %>%
  ungroup()

# Convert 'fye' to Date format
return_national_fye <- return_national_fye %>%
  mutate(fye = as.Date(fye))

return_10yr_filtered <- return_national_fye %>%
  filter(fye >= as.Date("2013-01-01") & fye <= as.Date("2022-12-31"))

return_20yr_filtered <- return_national_fye %>%
  filter(fye >= as.Date("2002-01-01") & fye <= as.Date("2022-12-31"))

# Calculate geometric means by fiscal year-end month
overall_geo_avg_10yr <- return_10yr_filtered %>%
  group_by(month) %>%
  summarise(geo_avg_10yr = prod(1 + national_return)^(1 / n()) - 1, .groups = "drop")

overall_geo_avg_20yr <- return_20yr_filtered %>%
  group_by(month) %>%
  summarise(geo_avg_20yr = prod(1 + national_return)^(1 / n()) - 1, .groups = "drop")

# Calculate overall geometric averages across all months
overall_geo_avg_10yr_all <- prod(1 + overall_geo_avg_10yr$geo_avg_10yr)^(1 / nrow(overall_geo_avg_10yr)) - 1
overall_geo_avg_20yr_all <- prod(1 + overall_geo_avg_20yr$geo_avg_20yr)^(1 / nrow(overall_geo_avg_20yr)) - 1

# Print the overall geometric averages
print(paste("Overall 10-Year Geometric Average Return (2013-2022):", round(overall_geo_avg_10yr_all * 100, 2), "%"))
print(paste("Overall 20-Year Geometric Average Return (2002-2022):", round(overall_geo_avg_20yr_all * 100, 2), "%"))


```


# 3.2 Figure 3: 10 & 20-Year Average Public Pension Returns By Fiscal Year End Month

```{r}

plot_20yr <- ggplot(overall_geo_avg_20yr, aes(x = as.factor(month), y = geo_avg_20yr)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_hline(yintercept = overall_geo_avg_20yr_all, color = "black", linetype = "dashed", size = 1) +  # Add overall geometric average line
  annotate("text", x = Inf, y = overall_geo_avg_20yr_all, label = paste("Avg:", round(overall_geo_avg_20yr_all * 100, 2), "%"), 
           vjust = -0.5, hjust = 1.2, size = 5, color = "black", fontface = "bold") +  # Add text for overall geometric average
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Month", y = "20-Year Average Return (2002-2022)", title = "20-Year Average Return by Month") +
  theme_reason()

# Plot for 10-year average returns by month with overall geometric average line and label
plot_10yr <- ggplot(overall_geo_avg_10yr, aes(x = as.factor(month), y = geo_avg_10yr)) +
  geom_bar(stat = "identity", fill = "orange") +
  geom_hline(yintercept = overall_geo_avg_10yr_all, color = "black", linetype = "dashed", size = 1) +  # Add overall geometric average line
  annotate("text", x = Inf, y = overall_geo_avg_10yr_all, label = paste("Avg:", round(overall_geo_avg_10yr_all * 100, 2), "%"), 
           vjust = -0.5, hjust = 1.2, size = 5, color = "black", fontface = "bold") +  # Add text for overall geometric average
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = reason_colors) + 
  labs(x = "Month", y = "10-Year Average Return (2013-2022)", title = "10-Year Average Return by Month") +
  theme_reason()

# Arrange the plots in a grid
grid.arrange(plot_20yr, plot_10yr, ncol = 2)
```

```{r}

plot_20yr2 <- ggplot(overall_geo_avg_20yr, aes(x = as.factor(month), y = geo_avg_20yr)) +
  geom_bar(stat = "identity", fill = "#FF6733") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(x = "Month", y = "20-Year Average Return (2002-2022)", title = "20-Year Average Return by Month") +
  theme_reason() 

# Plot for 10-year average returns by month with overall geometric average line and label
plot_10yr2 <- ggplot(overall_geo_avg_10yr, aes(x = as.factor(month), y = geo_avg_10yr)) +
  geom_bar(stat = "identity", fill = "#2F456B") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_manual(values = reason_colors) + 
  labs(x = "Month", y = "10-Year Average Return (2013-2022)", title = "10-Year Average Return by Month") +
  theme_reason() 


# Arrange the plots in a grid
G3 <- grid.arrange(plot_20yr2, plot_10yr2, ncol = 2)
```

# 3.3 Average 10 and 20 year Returns by Fiscal Year End Month Cohort   

```{r}
average_returns_summary <- overall_geo_avg_10yr %>%  # Changed to geometric averages
  rename("10-Year Avg Return" = geo_avg_10yr) %>%
  left_join(overall_geo_avg_20yr %>% rename("20-Year Avg Return" = geo_avg_20yr), by = "month") %>%
  mutate(
    `10-Year Avg Return` = round(`10-Year Avg Return` * 100, 2),
    `20-Year Avg Return` = round(`20-Year Avg Return` * 100, 2)
  )

# Display the combined table with fiscal year-end month, 10-year and 20-year average returns
average_returns_summary %>%
  arrange(month) %>%
  knitr::kable(col.names = c("Fiscal Year-End Month", "10-Year Average Return (%)", "20-Year Average Return (%)"),
               format = "html",
               caption = "Average Returns per Fiscal Year-End Month (10-Year and 20-Year Basis)") %>%
  kableExtra::kable_styling("striped", full_width = F)  # Optional styling for better visualization


```


